---
title: "Functional Data Analysis of EEG Data Comparing Alcoholics & Controls"
author: "Aoife O'Brien & Alannah Blacoe"
output:
  revealjs::revealjs_presentation:
    self_contained: true
    theme: white
    transition: fade
    incremental: true
    css: style.css

    
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
library(patchwork)
library(fdapace)
library(fda)
library(tibble)

firsttrial <- readRDS("firsttrial_data.rds")     
fp1_s2nomatch_control <- readRDS("fp1_control_fd.rds")
fp1_s2nomatch_alcohol <- readRDS("fp1_alcohol_fd.rds")
t_test_fp1_s2nomatch <- tperm.fd(fp1_s2nomatch_control, fp1_s2nomatch_alcohol)
fpca_scores_fourier <- readRDS("fpca_scores_fourier.rds")  
final_clusters <- readRDS("final_clusters.rds")    
res.fp1 <- readRDS("res.fp1.rds")                  
res.fp2 <- readRDS("res.fp2.rds") 
results <- readRDS("nharm_tuning_results.rds")
```

## Objective

* Investigate EEG correlates of genetic predisposition to alcoholism

* Use Functional Data Analysis (FDA) to compare alcoholic vs control EEG patterns

* Explore clustering and prediction of group membership from EEG curves

---

## Background & Motivation

* EEG produces hundreds of measurements per second.
* Traditional stats simplify this to averages — losing shape.
* FDA treats EEG as **curves over time**.
* FPCA extracts dominant patterns from these curves.

---

## The Dataset

* **Subjects:** Alcoholic & Control groups  
* **Electrodes:** 64 scalp sites (256 Hz sampling)  
* **Focus:** FP1 and FP2 frontal electrodes  
* **Stimuli:**  
  - S1 Object (single image)  
  - S2 Match (same second image)  
  - S2 Non-match (different second image)

---

## Research Questions

1. Do alcoholic and control participants show different EEG response patterns?  
2. How do visual stimuli impact EEG curves?  
3. Can we group participants based on EEG shape using clustering?  
4. Can we predict group membership from EEG data?

---

## Why Functional Data?

* EEG curves are not points — they're shapes.

* FDA handles entire time-dependent signal

* FPCA captures shared shapes across participants


## Raw EEG Curve from FP1
```{r, echo=FALSE, fig.align='center', out.width='70%'}
example_subject <- firsttrial %>% filter(sensor.position == "FP1") %>% distinct(name) %>% slice(1) %>% pull(name)
eeg_example <- firsttrial %>% filter(name == example_subject, sensor.position == "FP1")
ggplot(eeg_example, aes(x = time, y = sensor.value)) +
  geom_line(colour = "#00BFC4") +
  labs(title = paste("EEG Curve for", example_subject), x = "Time (s)", y = "Sensor Value") +
  theme_minimal()
```


## Permutation Test 


- A **non-parametric test** used to compare groups  
- Shuffles group labels (e.g., alcoholic vs control) many times  
- Builds a **null distribution** of test statistics  
- Compares actual observed stat to that distribution  
- No assumptions about normality — perfect for complex data like EEG

---

## FP1 Electrode (S2 Non-match)
```{r, echo=FALSE, fig.align='center'}
t_test_fp1_s2nomatch <- tperm.fd(fp1_s2nomatch_control, fp1_s2nomatch_alcohol)

```


---

## FPCA Model

\[
X(t) = \mu(t) + \sum_{k=1}^{K} \xi_k \phi_k(t)
\]

* \(\mu(t)\): mean EEG curve  
* \(\phi_k(t)\): principal components (shapes)  
* \(\xi_k\): subject-level scores

---

## FPCA Summary – FP1

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(patchwork)

# Scree Plot
varprop <- res.fp1$varprop
scree_df <- tibble(PC = paste0("PC", seq_along(varprop)), Variance = varprop)
p1 <- scree_df %>%
  slice(1:10) %>%
  ggplot(aes(x = PC, y = Variance)) +
  geom_col(fill = "#00BFC4") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Scree Plot (FP1)", x = "Component", y = "Variance Explained") +
  theme_minimal()

# Mean EEG Curve
times <- seq(0, 1, length.out = 256)
mean_vals <- as.vector(eval.fd(times, res.fp1$meanfd))
mean_df <- tibble(time = times, amplitude = mean_vals)
p2 <- ggplot(mean_df, aes(x = time, y = amplitude)) +
  geom_line(color = "black") +
  labs(title = "Mean EEG Curve (FP1)", x = "Time (s)", y = "Amplitude") +
  theme_minimal()

# Components 1–3
harmonics_df <- as_tibble(eval.fd(times, res.fp1$harmonics[1:3]))
names(harmonics_df) <- c("Component 1", "Component 2", "Component 3")
harmonics_df$time <- times
harmonics_long <- pivot_longer(harmonics_df, cols = starts_with("Component"),
                                names_to = "Component", values_to = "Amplitude")

p3 <- ggplot(harmonics_long, aes(x = time, y = Amplitude, colour = Component, linetype = Component)) +
  geom_line(size = 1) +
  scale_colour_manual(values = c("black", "#00BFC4", "#F8766D")) +
  labs(title = "Components 1–3 (FP1)", x = "Time (s)", y = "Amplitude") +
  theme_minimal() +
  theme(legend.position = "right")

# Layout: Scree + Mean on top, Components wide on bottom
(p1 | p2) / p3

```


---

## RQ1: Do EEG Patterns Differ by Group?

* Permutation test showed significant group differences
* Strongest for **non-matching stimuli** (S2 Non-match)

---

## RQ2: Effect of Stimuli on EEG

```{r, echo=FALSE, fig.align='center'}
ggplot(fpca_scores_fourier, aes(x = matching.condition, y = FP1.1, fill = matching.condition)) +
  geom_boxplot() +
  labs(title = "FP1 Component 1 by Stimulus", x = "Stimulus", y = "FP1 - PC1 Score") +
  theme_minimal()
```

---

## RQ3: Clustering on EEG Shape

```{r, echo=FALSE, fig.align='center'}
ggplot(final_clusters, aes(x = FP1_1, y = FP2_1, colour = cluster)) +
  geom_point(size = 3) +
  labs(title = "K-means Clustering (3 groups)", x = "FP1 PC1", y = "FP2 PC1") +
  theme_minimal()
```


---

## RQ4: Can We Predict Group from EEG?

* Models: Random Forest, Multinomial
* Inputs: PC1–PC6 from FP1 & FP2
* Compared **Fourier** vs **B-spline** basis

---


```{r, echo=FALSE, fig.align='center'}
ggplot(results, aes(x = NHarm, y = Matching_RF, colour = Basis)) +
  geom_line(size = 1) +
  geom_point() +
  labs(title = "RF Accuracy for Matching Condition by Basis Type",
       x = "Number of Components", y = "Accuracy") +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal()
```
---


```{r, echo=FALSE, fig.align='center'}
library(tidyr)

results_long <- results %>%
  pivot_longer(cols = c(Matching_RF, Subject_RF),
               names_to = "Task", values_to = "Accuracy")

ggplot(results_long, aes(x = NHarm, y = Accuracy, colour = Basis)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~Task, ncol = 1) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Random Forest Accuracy by Basis Type",
    x = "Number of Components (NHarm)",
    y = "Accuracy"
  ) +
  theme_minimal()

```
---

## Component Tuning Results

```{r, echo=FALSE, fig.align='center'}
results <- readRDS("nharm_tuning_results.rds")
results_long <- results %>%
  pivot_longer(cols = starts_with("Matching_"), names_to = "Model_Matching", values_to = "Accuracy") %>%
  mutate(Model = gsub("Matching_", "", Model_Matching), Config = paste(Basis, NHarm, sep = "-")) %>%
  arrange(desc(Accuracy)) %>%
  slice_head(n = 5)

ggplot(results_long, aes(x = reorder(Config, -Accuracy), y = Accuracy, fill = Model)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(title = "Top 5 FPCA Configurations for Stimulus Prediction", x = "Basis-Type + Components", y = "Accuracy", fill = "Model") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

---

## Final Summary

* EEG curves differ by subject group and stimulus
* FPCA reduced complexity — top 3 components explained >90% variance
* Random Forest with **Fourier basis** performed best (75% accuracy)

---

## Limitations 

* Small sample size (only 16 subjects)

* Limited to FP1 and FP2 sensors

* Only the first trial per condition included

* Possible subject-level confounding not accounted for

* Clustering and modelling may overfit due to high dimensionality
---

## Conclusion & Next Steps

**Conclusion**
- FDA captured group-level EEG differences
- FPCA scores worked well as predictors

**Next Steps**
- Extend to all 64 electrodes
- Explore time-locked responses (e.g., ERP)
- Try deep learning models (e.g., CNN)

---

## Questions?
