---
title: "2_plots"
output: html_document
date: "2025-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(fda.usc)
```

```{r}
firsttrial <- readRDS("firsttrial_data.rds")

head(firsttrial)
summary(firsttrial)
```


### Individual plots comparing each sensor position and matching condition for Alcoholics and controls (Cleaned data)

```{r}
library(ggplot2)
library(dplyr)

combined_data <- firsttrial %>%
  filter((sensor.position %in% c("FP1", "FP2")) &
           (matching.condition %in% c("S1 obj", "S2 match", "S2 nomatch,")) &
           (subject.identifier %in% c("c", "a"))) %>%
  mutate(group = ifelse(subject.identifier == "c", "Control", "Alcoholic")) %>%
 dplyr::  select(sensor.position, matching.condition, group, time, sensor.value)

unique_conditions <- unique(combined_data$matching.condition)
unique_positions <- unique(combined_data$sensor.position)

for (condition in unique_conditions) {
  for (position in unique_positions) {
    plot_data <- combined_data %>%
      filter(matching.condition == condition, sensor.position == position)
    
    plot_title <- paste("Sensor Values: ", position, " - ", condition, sep = "")
    
    plot <- ggplot(plot_data, aes(x = time, y = sensor.value, color = group)) +
      geom_line() +
      labs(
        title = plot_title,
        x = "Time (ms)",
        y = "Sensor Value",
        color = "Group"
      ) +
      theme_minimal()
    
    print(plot)
  }
}

```

###Plots comparing each sensor position and matching condition for Alcoholics and controls (Cleaned data)

```{r}
table(firsttrial$sensor.position, firsttrial$matching.condition)
library(ggplot2)
library(dplyr)

combined_data <- firsttrial %>%
  filter(
    sensor.position %in% c("FP1", "FP2"),  
    matching.condition %in% c("S1 obj", "S2 match", "S2 nomatch,")  
  ) %>%
  mutate(
    group = ifelse(subject.identifier == "c", "Control", "Alcoholic"),
    matching.condition = factor(
      matching.condition,
      levels = c("S1 obj", "S2 match", "S2 nomatch,")  
    )
  ) %>%
  dplyr:: select(sensor.position, matching.condition, group, time, sensor.value)


ggplot(combined_data, aes(x = time, y = sensor.value, color = group)) +
  geom_line() +
  facet_grid(sensor.position ~ matching.condition, scales = "free_y", drop = FALSE) +  
  labs(
    title = "Sensor Values Over Time",
    subtitle = "Comparing Control and Alcoholic Groups Across Conditions and Positions",
    x = "Time (ms)",
    y = "Sensor Value",
    color = "Group"
  ) +
  theme_minimal()

```

### Plots comparing each sensor position and matching condition for Alcoholics and controls (Uncleaned data)

```{r}
table(dat$sensor.position, dat$matching.condition)
library(ggplot2)
library(dplyr)


combined_data <- dat %>%
  filter(
    sensor.position %in% c("FP1", "FP2"),  
    matching.condition %in% c("S1 obj", "S2 match", "S2 nomatch,")  
  ) %>%
  mutate(
    group = ifelse(subject.identifier == "c", "Control", "Alcoholic"),
    matching.condition = factor(
      matching.condition,
      levels = c("S1 obj", "S2 match", "S2 nomatch,")  
    )
  ) %>%
  dplyr:: select(sensor.position, matching.condition, group, time, sensor.value)


ggplot(combined_data, aes(x = time, y = sensor.value, color = group)) +
  geom_line() +
  facet_grid(sensor.position ~ matching.condition, scales = "free_y", drop = FALSE) + 
  labs(
    title = "Sensor Values Over Time",
    subtitle = "Comparing Control and Alcoholic Groups Across Conditions and Positions(Uncleaned)",
    x = "Time (ms)",
    y = "Sensor Value",
    color = "Group"
  ) +
  theme_minimal()

```

### Pointwise Differences of Stimuli on Alcoholics and Controls

```{r}

control_s1 <- firsttrial %>%
  filter(sensor.position == "FP1",
         matching.condition == "S1 obj",
         subject.identifier == "c") %>%
  dplyr::select(name, time, sensor.value)
control_fd_s1_matrix <- matrix(as.numeric(control_s1$sensor.value), 
                               nrow = 256, ncol = 8)


alcohol_s1 <- firsttrial %>%
  filter(sensor.position == "FP1",
         matching.condition == "S1 obj",
         subject.identifier == "a") %>%
  dplyr::select(name, time, sensor.value)
alcohol_fd_s1_matrix <- matrix(as.numeric(alcohol_s1$sensor.value), 
                               nrow = 256, ncol = 8)


control_s2match <- firsttrial %>%
  filter(sensor.position == "FP1",
         matching.condition == "S2 match",
         subject.identifier == "c") %>%
  dplyr::select(name, time, sensor.value)
control_fd_s2match_matrix <- matrix(as.numeric(control_s2match$sensor.value), 
                                    nrow = 256, ncol = 8)


alcohol_s2match <- firsttrial %>%
  filter(sensor.position == "FP1",
         matching.condition == "S2 match",
         subject.identifier == "a") %>%
  dplyr::select(name, time, sensor.value)
alcohol_fd_s2match_matrix <- matrix(as.numeric(alcohol_s2match$sensor.value), 
                                    nrow = 256, ncol = 8)


control_s2nomatch <- firsttrial %>%
  filter(sensor.position == "FP1",
         matching.condition == "S2 nomatch,",
         subject.identifier == "c") %>%
  dplyr::select(name, time, sensor.value)
control_fd_s2nomatch_matrix <- matrix(as.numeric(control_s2nomatch$sensor.value), 
                                      nrow = 256, ncol = 8)


alcohol_s2nomatch <- firsttrial %>%
  filter(sensor.position == "FP1",
         matching.condition == "S2 nomatch,",
         subject.identifier == "a") %>%
  dplyr::select(name, time, sensor.value)
alcohol_fd_s2nomatch_matrix <- matrix(as.numeric(alcohol_s2nomatch$sensor.value), 
                                      nrow = 256, ncol = 8)


time <- seq(0, 1, length.out = 256)
norder <- 10
nbasis <- length(time) + norder - 2
basis <- create.bspline.basis(range(time), nbasis, norder, time)

Lfdobj <- 4
lambda <- 1e-2
fdParObj <- fdPar(fd(matrix(0, nbasis, 1), basis), Lfdobj, lambda)


control_fd_s1 <- smooth.basis(time, control_fd_s1_matrix, fdParObj)$fd
alcohol_fd_s1 <- smooth.basis(time, alcohol_fd_s1_matrix, fdParObj)$fd
control_fd_s2match <- smooth.basis(time, control_fd_s2match_matrix, fdParObj)$fd
alcohol_fd_s2match <- smooth.basis(time, alcohol_fd_s2match_matrix, fdParObj)$fd
control_fd_s2nomatch <- smooth.basis(time, control_fd_s2nomatch_matrix, fdParObj)$fd
alcohol_fd_s2nomatch <- smooth.basis(time, alcohol_fd_s2nomatch_matrix, fdParObj)$fd


mean_control_s1 <- mean.fd(control_fd_s1)
mean_alcohol_s1 <- mean.fd(alcohol_fd_s1)
mean_control_s2match <- mean.fd(control_fd_s2match)
mean_alcohol_s2match <- mean.fd(alcohol_fd_s2match)
mean_control_s2nomatch <- mean.fd(control_fd_s2nomatch)
mean_alcohol_s2nomatch <- mean.fd(alcohol_fd_s2nomatch)


diff_s1 <- mean_alcohol_s1 - mean_control_s1
diff_s2match <- mean_alcohol_s2match - mean_control_s2match
diff_s2nomatch <- mean_alcohol_s2nomatch - mean_control_s2nomatch


y_range <- range(c(diff_s1$coefs, diff_s2match$coefs, diff_s2nomatch$coefs), finite = TRUE)


plot(diff_s1, col = "blue", lwd = 2, lty = 1, 
     main = "Pointwise Differences of Stimuli on Alcoholics and Controls",
     ylab = "Difference in Sensor Value", xlab = "Time", ylim = y_range)
lines(diff_s2match, col = "purple", lwd = 2, lty = 2)
lines(diff_s2nomatch, col = "red", lwd = 2, lty = 3)
abline(h = 0, col = "black", lty = 3)  


legend("topright", legend = c("S1", "S2 Match", "S2 No Match"),
       col = c("blue", "purple", "red"), lty = c(1, 2, 3), lwd = 2)


```

### Pointwise Differences Between FP1 and FP2

```{r}

alcohol_fp1 <- firsttrial %>%
  filter(sensor.position == "FP1",
         subject.identifier == "a") %>%
  dplyr::select(name, time, sensor.value)
alcohol_fp1_matrix <- matrix(as.numeric(alcohol_fp1$sensor.value), 
                             nrow = 256, ncol = 8) 

control_fp1 <- firsttrial %>%
  filter(sensor.position == "FP1",
         subject.identifier == "c") %>%
  dplyr::select(name, time, sensor.value)
control_fp1_matrix <- matrix(as.numeric(control_fp1$sensor.value), 
                             nrow = 256, ncol = 8)  


alcohol_fp2 <- firsttrial %>%
  filter(sensor.position == "FP2",
         subject.identifier == "a") %>%
  dplyr::select(name, time, sensor.value)
alcohol_fp2_matrix <- matrix(as.numeric(alcohol_fp2$sensor.value), 
                             nrow = 256, ncol = 8)  


control_fp2 <- firsttrial %>%
  filter(sensor.position == "FP2",
         subject.identifier == "c") %>%
  dplyr::select(name, time, sensor.value)
control_fp2_matrix <- matrix(as.numeric(control_fp2$sensor.value), 
                             nrow = 256, ncol = 8)  


time <- seq(0, 1, length.out = 256)
norder <- 10
nbasis <- length(time) + norder - 2
basis <- create.bspline.basis(range(time), nbasis, norder, time)


Lfdobj <- 4
lambda <- 1e-2
fdParObj <- fdPar(fd(matrix(0, nbasis, 1), basis), Lfdobj, lambda)


alcohol_fp1_fd <- smooth.basis(time, alcohol_fp1_matrix, fdParObj)$fd
control_fp1_fd <- smooth.basis(time, control_fp1_matrix, fdParObj)$fd


alcohol_fp2_fd <- smooth.basis(time, alcohol_fp2_matrix, fdParObj)$fd
control_fp2_fd <- smooth.basis(time, control_fp2_matrix, fdParObj)$fd


mean_alcohol_fp1 <- mean.fd(alcohol_fp1_fd)
mean_alcohol_fp2 <- mean.fd(alcohol_fp2_fd)
mean_control_fp1 <- mean.fd(control_fp1_fd)
mean_control_fp2 <- mean.fd(control_fp2_fd)


mean_diff_alcohol <- mean_alcohol_fp1 - mean_alcohol_fp2
mean_diff_control <- mean_control_fp1 - mean_control_fp2


plot(mean_diff_alcohol, col = "red", lwd = 2, 
     main = "Pointwise Differences Between FP1 and FP2",
     ylab = "Difference in Sensor Value", xlab = "Time", ylim = c(-2, 6))
lines(mean_diff_control, col = "blue", lwd = 2, lty = 2)
legend("topright", legend = c("Alcoholics", "Controls"),
       col = c("red", "blue"), lty = c(1, 2), lwd = 2)

```


```{r}
# Plot a single EEG curve (FP1) from one subject
example_subject <- firsttrial %>%
  filter(sensor.position == "FP1") %>%
  distinct(name) %>%
  slice(1) %>% pull(name)

eeg_example <- firsttrial %>%
  filter(name == example_subject, sensor.position == "FP1")

ggplot(eeg_example, aes(x = time, y = sensor.value)) +
  geom_line(colour = "steelblue", size = 1) +
  labs(title = paste("EEG Curve for", example_subject, "(FP1)"),
       x = "Time (seconds)", y = "Sensor Value") +
  theme_minimal()

```

```{r}
# Create a dummy tibble if you don't have real results yet
model_results <- tibble::tibble(
  Model = c("Fourier + RF", "Fourier + Lasso", 
            "B-spline + RF", "B-spline + Lasso"),
  Accuracy = c(0.88, 0.83, 0.81, 0.78)  # Replace with your real scores
)

# Bar plot of classification accuracy
ggplot(model_results, aes(x = Model, y = Accuracy, fill = Model)) +
  geom_col(width = 0.6) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(title = "Model Accuracy by Method & Basis",
       x = "Model",
       y = "Accuracy") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 30, hjust = 1))

```
```{r}
# Pick a subject
subject_name <- unique(firsttrial$name)[1]  # or change to a specific one

# Filter FP1 data for the selected subject
subject_data <- firsttrial %>%
  filter(sensor.position == "FP1", name == subject_name) %>%
  arrange(time)

# Extract time and raw signal
time <- subject_data$time
raw <- subject_data$sensor.value

# Evaluate fits (force to numeric in case of 1-column matrix)
bspline_fit <- as.numeric(eval.fd(time, res_fp1$meanfd + 2 * res_fp1$harmonics[1]))
fourier_fit <- as.numeric(eval.fd(time, res.fp1$meanfd + 2 * res.fp1$harmonics[1]))

# Build a tidy dataframe
plot_df <- data.frame(
  time = time,
  raw = raw,
  bspline = bspline_fit,
  fourier = fourier_fit
)

# Plot
library(ggplot2)

ggplot(plot_df, aes(x = time)) +
  geom_line(aes(y = raw), color = "gray70", alpha = 0.7, size = 0.5) +
  geom_line(aes(y = bspline), color = "red", size = 1) +
  geom_line(aes(y = fourier), color = "blue", size = 1) +
  labs(
    title = paste("EEG Curve Smoothing –", subject_name),
    subtitle = "B-spline (red) vs Fourier (blue) vs Raw (gray)",
    y = "EEG Value",
    x = "Time (seconds)"
  ) +
  theme_minimal()

```


```{r}
library(caret)
# Confusion matrix for LASSO
confusionMatrix(data = factor(pred_lasso), reference = factor(truth))

# Confusion matrix for Random Forest
confusionMatrix(data = factor(pred_rf), reference = factor(truth))

```

```{r}
library(ggplot2)
library(dplyr)

cm_df <- data.frame(
  Model = rep(c("RF", "LASSO"), each = 4),
  Reference = rep(c("Alcoholic", "Control"), 4),
  Prediction = rep(c("Alcoholic", "Control"), times = 4),
  Count = c(10, 2, 3, 13, 9, 4, 5, 12)  # Replace with your own confusion matrix counts
)

ggplot(cm_df, aes(x = Reference, y = Prediction, fill = Count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkred") +
  facet_wrap(~Model) +
  labs(title = "Confusion Matrices", fill = "Count") +
  theme_minimal()

```

```{r}
metrics_df <- data.frame(
  Model = rep(c("Lasso", "Random Forest"), each = 3),
  Basis = rep(c("B-spline", "Fourier"), times = 3),
  Metric = rep(c("Accuracy", "Sensitivity", "Specificity"), 2),
  Value = c(0.78, 0.75, 0.80, 0.84, 0.81, 0.85)  # Replace with your values
)

```


```{r}
ggplot(metrics_df, aes(x = Model, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Basis) +
  ylim(0, 1) +
  labs(title = "Model Performance Comparison",
       y = "Metric Value",
       x = "Model") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal()

```


```{r}
library(dplyr)
library(ggplot2)

# Create group variable: Alcoholic vs Control
firsttrial <- firsttrial %>%
  mutate(group = ifelse(grepl("^a", subject.identifier), "Alcoholic", "Control"))

# Keep only FP1 (or change to FP2 if you want that instead)
eeg_mean_df <- firsttrial %>%
  filter(sensor.position == "FP1") %>%
  group_by(group, matching.condition, time) %>%
  summarise(
    mean_value = mean(sensor.value, na.rm = TRUE),
    sd_value = sd(sensor.value, na.rm = TRUE),
    .groups = "drop"
  )

# Plot: Mean curves with confidence band (optional)
ggplot(eeg_mean_df, aes(x = time, y = mean_value, color = group, fill = group)) +
  geom_line(size = 1.1) +
  geom_ribbon(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value), alpha = 0.15, color = NA) +
  facet_wrap(~matching.condition) +
  labs(
    title = "Mean EEG Curves by Group and Condition (FP1)",
    x = "Time (seconds)",
    y = "EEG Voltage (uV)",   # ← updated from µV to uV
    color = "Group",
    fill = "Group"
  ) +
  theme_minimal(base_size = 14)

```
```{r}
# Example variables — replace with your actual results
# null_dist = vector of permuted test stats
# obs_stat = observed test stat

# For illustration — replace these with your own
# (e.g., S2_NoMatch_perm_results$res_fp1 or however it's stored)
null_dist <- rnorm(1000, mean = 0, sd = 1)  # simulated null
obs_stat <- 2.5                             # your actual observed value

# Plot
library(ggplot2)

ggplot(data.frame(null = null_dist), aes(x = null)) +
  geom_histogram(aes(y = ..density..), bins = 50, fill = "gray80", color = "gray50") +
  geom_density(color = "black", size = 1) +
  geom_vline(xintercept = obs_stat, color = "red", size = 1.2, linetype = "dashed") +
  labs(
    title = "Permutation Test – S2 No-Match (FP1)",
    subtitle = "Observed test statistic vs null distribution",
    x = "Test Statistic (Area Between Group Curves)",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)

```

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)

# Select and reshape PC1–PC3 scores along with matching condition
score_data <- fpca_scores_fourier %>%
  select(name, matching.condition, starts_with("FP1."), starts_with("FP2.")) %>%
  pivot_longer(cols = c(FP1.1, FP1.2, FP1.3), names_to = "PC", values_to = "Score")

# Optional: clean up PC labels
score_data$PC <- factor(score_data$PC, labels = c("PC1", "PC2", "PC3"))

# Plot for FP1 scores across stimuli
ggplot(score_data, aes(x = matching.condition, y = Score, fill = PC)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
  facet_wrap(~PC, scales = "free_y") +
  labs(
    title = "FP1 – FPCA Scores by Stimulus Condition",
    x = "Stimulus Type",
    y = "FPCA Score",
    fill = "Component"
  ) +
  theme_minimal(base_size = 14) +
  scale_fill_brewer(palette = "Set2")

```

```{r}
library(fda)
library(dplyr)
library(ggplot2)

# Filter to FP1
fp1_data <- firsttrial %>%
  filter(sensor.position == "FP1", !is.na(matching.condition)) %>%
  mutate(stimulus = case_when(
    grepl("S1", matching.condition, ignore.case = TRUE) ~ "S1 Object",
    grepl("S2 match", matching.condition, ignore.case = TRUE) ~ "S2 Match",
    grepl("S2 nomatch", matching.condition, ignore.case = TRUE) ~ "S2 No-Match",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(stimulus))

# Average per time point and stimulus
avg_data <- fp1_data %>%
  group_by(stimulus, time) %>%
  summarise(mean_val = mean(sensor.value, na.rm = TRUE), .groups = "drop")

# Smooth using Fourier basis
basis <- create.fourier.basis(c(0, 1), nbasis = 15)

smoothed_df_list <- avg_data %>%
  group_split(stimulus) %>%
  lapply(function(df) {
    smooth_fd <- smooth.basis(df$time, df$mean_val, basis)$fd
    smoothed_vals <- eval.fd(seq(0, 1, length.out = 256), smooth_fd)
    data.frame(
      time = seq(0, 1, length.out = 256),
      value = smoothed_vals,
      stimulus = unique(df$stimulus)
    )
  })

plot_data <- bind_rows(smoothed_df_list)

# Plot
ggplot(plot_data, aes(x = time, y = value, color = stimulus, linetype = stimulus)) +
  geom_line(size = 1.3) +
  scale_color_manual(values = c("blue", "purple", "red")) +
  scale_linetype_manual(values = c("solid", "dashed", "solid")) +
  labs(
    title = "Smoothed Mean EEG Curves by Stimulus (FP1)",
    x = "Time (seconds)",
    y = "EEG Voltage (μV)",
    color = "Stimulus Type",
    linetype = "Stimulus Type"
  ) +
  theme_minimal(base_size = 14)

```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# Assuming `fpca_scores_bspline` contains PC1, PC2, and stimulus info
df <- fpca_scores_bspline %>%
  select(name, PC1 = FP1.1, PC2 = FP1.2, Stimulus = matching.condition) %>%
  pivot_longer(cols = c(PC1, PC2), names_to = "Component", values_to = "Score")

ggplot(df, aes(x = Stimulus, y = Score, fill = Component)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6) +
  facet_wrap(~Component, scales = "free_y") +
  labs(title = "FP1 – FPCA Scores by Stimulus Condition",
       x = "Stimulus Type", y = "FPCA Score") +
  theme_minimal(base_size = 14) +
  scale_fill_brewer(palette = "Set2")

```

```{r}
library(dplyr)
library(ggplot2)

# Create Group label from subject.identifier
final_clusters <- final_clusters %>%
  mutate(Group = ifelse(grepl("a", subject.identifier), "Alcoholic", "Control"))

# Optionally shorten stimulus labels
final_clusters <- final_clusters %>%
  mutate(Stimulus = case_when(
    matching.condition == "S1 obj" ~ "S1",
    matching.condition == "S2 match" ~ "S2 Match",
    matching.condition == "S2 nomatch" ~ "S2 No-Match",
    TRUE ~ matching.condition
  ))
# Bar plot: % Alcoholic vs Control per cluster
ggplot(final_clusters, aes(x = as.factor(cluster), fill = Group)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Cluster Composition by Group",
    x = "Cluster",
    y = "Proportion",
    fill = "Group"
  ) +
  theme_minimal(base_size = 14)


```
```{r}
# Bar plot: Stimulus breakdown per cluster
ggplot(final_clusters, aes(x = as.factor(cluster), fill = Stimulus)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Cluster Composition by Stimulus Type",
    x = "Cluster",
    y = "Proportion",
    fill = "Stimulus"
  ) +
  theme_minimal(base_size = 14)

```

new *** 

```{r}
# Plot eigenfunctions for FP1 and FP2 using Fourier-based FPCA
plot.pca.fd(res.fp1, nx = 256, harm = 1:3, expand = 0.05)
plot.pca.fd(res.fp2, nx = 256, harm = 1:3, expand = 0.05)

```

```{r}
# Make sure PC1 and PC2 are in the data
library(ggplot2)
ggplot(final_clusters, aes(x = FP1.1, y = FP1.2, colour = factor(cluster))) +
  geom_point(alpha = 0.8, size = 2) +
  labs(
    title = "FPCA Feature Space: PC1 vs PC2 (FP1 Scores)",
    x = "Principal Component 1 (FP1.1)",
    y = "Principal Component 2 (FP1.2)",
    colour = "Cluster"
  ) +
  theme_minimal(base_size = 14)

```


```{r}
ggplot(final_clusters, aes(x = FP1.1, y = FP1.2, colour = factor(cluster), shape = Group)) +
  geom_point(alpha = 0.8, size = 2) +
  facet_wrap(~ Stimulus) +
  labs(
    title = "PC1 vs PC2 Scatter: Clusters by Group and Stimulus",
    x = "Principal Component 1 (FP1.1)",
    y = "Principal Component 2 (FP1.2)",
    colour = "Cluster",
    shape = "Group"
  ) +
  theme_minimal(base_size = 14)

```

```{r}
library(tidyr)
library(dplyr)

results_long <- results %>%
  pivot_longer(
    cols = starts_with("Matching") | starts_with("Subject"),
    names_to = c("Task", "Model"),
    names_sep = "_",
    values_to = "Accuracy"
  )
```

```{r}
library(plotly)

# Interactive line plot comparing model performance
plot_ly(results_long, 
        x = ~NHarm, 
        y = ~Accuracy, 
        color = ~Model, 
        linetype = ~Task,
        type = 'scatter', 
        mode = 'lines+markers',
        split = ~interaction(Task, Model),
        hoverinfo = 'text',
        text = ~paste("Basis:", Basis,
                      "<br>Task:", Task,
                      "<br>Model:", Model,
                      "<br>NHarm:", NHarm,
                      "<br>Accuracy:", round(Accuracy, 3))) %>%
  layout(title = "Model Accuracy by Number of Harmonics",
         xaxis = list(title = "Number of Harmonics"),
         yaxis = list(title = "Accuracy"),
         legend = list(title = list(text = "Model Type")),
         hovermode = "closest")

```
```{r}
plot_ly(results_long, 
        x = ~NHarm, 
        y = ~Accuracy, 
        color = ~Model, 
        linetype = ~Task,
        split = ~interaction(Task, Model),
        type = 'scatter',
        mode = 'lines+markers',
        hoverinfo = 'text',
        text = ~paste("Basis:", Basis,
                      "<br>Task:", Task,
                      "<br>Model:", Model,
                      "<br>NHarm:", NHarm,
                      "<br>Accuracy:", round(Accuracy, 3))) %>%
  layout(title = "Model Accuracy by NHarm (Faceted by Basis)",
         xaxis = list(title = "Number of Harmonics"),
         yaxis = list(title = "Accuracy"),
         legend = list(title = list(text = "Model Type")),
         hovermode = "closest") %>%
  subplot(nrows = 1, shareX = TRUE, shareY = TRUE) %>%
  add_trace(data = subset(results_long, Basis == "bspline"), name = "B-spline") %>%
  add_trace(data = subset(results_long, Basis == "fourier"), name = "Fourier")

```

