---
title: "4_pca_bspline"
output: html_document
date: "2025-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(fda.usc)
library(nnet) 
```

```{r}
firsttrial <- readRDS("firsttrial_data.rds")

head(firsttrial)
summary(firsttrial)
```


```{r}
# Define the time grid for 256 points between 0 and 1
time <- seq(0, 1, length.out = 256)

# Define B-spline basis parameters for smoothing FPCA data
norder <- 4
nbasis <-  length(time) / 2 
basis <- create.bspline.basis(rangeval = range(time), nbasis = nbasis, norder = norder)

# Define the penalty and smoothing parameter
Lfdobj <- 2    
lambda <- 1e-2
fdParObj <- fdPar(fd(matrix(0, nbasis, 1), basis), Lfdobj, lambda)

# Smoothing helper function: converts a sensor matrix into a functional data object.
smooth_fd <- function(sensor_matrix, time, fdParObj) {
  smooth.basis(time, sensor_matrix, fdParObj)$fd
}

# Extract FP1 sensor data
fp1_data <- firsttrial %>%
  filter(sensor.position == "FP1") %>%
  dplyr::select(name, matching.condition, subject.identifier, time, sensor.value) %>%
  arrange(matching.condition, subject.identifier, name, time)
```

# Don't pass time as breaks – it overloaded the basis function

# Too many breaks (256) violated the nbasis = norder + breaks - 2 rule

# Using default breaks avoids basis function mismatch

# This setup keeps the basis smoother and avoids overfitting

# Basis now has fewer functions → better smoothing for FPCA

# Clean fix: let fda package handle internal break placement

# Reduced complexity by letting basis use default knots



```{r}
# Extract FP1 ordering data at time 0 for modeling
fp_order <- firsttrial %>%
  filter(sensor.position == "FP1") %>%
  dplyr:: select(name, matching.condition, subject.identifier, time, sensor.value) %>%
  filter(time == 0) %>%
  arrange(matching.condition, subject.identifier, name)

# Extract the full FP1 time series data
fp1_data <- firsttrial %>%
  filter(sensor.position == "FP1") %>%
 dplyr::  select(name, matching.condition, subject.identifier, time, sensor.value) %>%
  arrange(matching.condition, subject.identifier, name, time)

```


```{r}
# Convert FP1 data to a matrix (256 x 48) and smooth it
pca_matrix_fp1 <- matrix(as.numeric(fp1_data$sensor.value), nrow = 256, ncol = 48)
pca_fd_fp1 <- smooth.basis(time, pca_matrix_fp1, fdParObj)$fd
plot(pca_fd_fp1)


nharm <- 7  # based on scree/eigenfunction plots
resfp1 <- pca.fd(pca_fd_fp1, nharm = nharm)
plot(resfp1)

saveRDS(resfp1,"resfp1.rds")

# Extract FPCA scores and prepare modeling data
scores <- data.frame(resfp1$scores)
colnames(scores) <- paste0("FP1.", seq_len(ncol(scores)))

# Create a factor for the condition based on subject.identifier
fp_order$condition_factor <- factor(fp_order$subject.identifier)

# Combine ordering information with FPCA scores
fpca_scores_bspline <- cbind(fp_order, scores)

```


```{r}
# Extract FP2 data
fp2_data <- firsttrial %>%
  filter(sensor.position == "FP2") %>%
  dplyr:: select(name, matching.condition, subject.identifier, time, sensor.value) %>%
  arrange(matching.condition, subject.identifier, name, time)

# Convert FP2 sensor data into a matrix (256 x 48) and smooth it
pca_matrix_fp2 <- matrix(as.numeric(fp2_data$sensor.value), nrow = 256, ncol = 48)
pca_fd_fp2 <- smooth.basis(time, pca_matrix_fp2, fdParObj)$fd

nharm <- 7  # based on scree/eigenfunction plots
resfp2 <- pca.fd(pca_fd_fp2, nharm = nharm)

saveRDS(resfp2,"resfp2.rds")

# Extract FP2 scores
scores2 <- data.frame(resfp2$scores)
colnames(scores2) <- paste0("FP2.",seq_len(ncol(scores2)))

```


```{r}
# Combine FP1 ordering data with FP1 and FP2 FPCA scores
fpca_scores_bspline <- cbind(fp_order, scores, scores2)

saveRDS(fpca_scores_bspline, "fpca_scores_bspline.rds")
```


```{r}
plot(pca_fd_fp1)
plot(pca_fd_fp2)
```

